---
title: "Homework 3"
author: "Bella Kormos"
format: html
---

[GitHub Repository](https://github.com/bellakormos/ENVS-193DS_homework-03) 
# Homework Set Up

Reading in packages:

```{r}
#| label: packages
#| message: false
library(tidyverse) # load in all packages
library(janitor)
library(here) # identifies computer's root folder
library(lubridate)

```

Reading in data:

```{r}
#| label: reading in data
#| message: false
salinity <- read_csv(here("data", "salinity-pickleweed.csv")) # store the data and save object as "salinity" 
my_data <- read_csv(here("data", "updated_personal-data.csv")) # Read in the personal data as an object named "my_data"

```

# Problem 1. Slough soil salinity

## a. An appropriate test
To determine the strength of the relationship between salinity and California pickleweed biomass, the two appropriate tests are the Pearson correlation and Spearman correlation tests. The pearson correlation (correlation coefficient, r) measures the strength between soil salinity and biomass, and assumes that variables are continuous, normally distributed, and have independent observations. The Spearman rank correlation is a non-parametric test that measures the stength of the monotonic relationship between ranked values of soil salinity and biomass, and it does not assume normality. 

## b. Create a visualization
```{r}
#| label: create visualizaiton
#| message: false
# base layer: ggplot
ggplot(data = salinity, # using salinity data frame
       aes(x = salinity_mS_cm, # soil salinity on x-axis (predictor)
           y = pickleweed)) + # pickleweed biomass on y-axis (response)
  # first layer: make scatter-plot 
  geom_point(color = "forestgreen",
             alpha = 0.7, # slight transparency 
             size = 2) + # slightly larger points
  # title plot, x-axis, y-axis
  labs(x = "Soil Salinity (mS/cm)", 
       y = "Pickleweed Biomass (g)",
       title = "Relationship between Soil Salinity and Pickleweed Biomass") +
  theme_light() # different from default theme
```


## c. Check your assumptions and run your test

### Assumption Checks

```{r}
#| label: assumption checks
#| message: false

# check normality of salinity variable using Shapiro-Wilk test
shapiro.test(salinity$salinity_mS_cm)

# check normality of biomass variable using Shapiro-Wilk test
shapiro.test(salinity$pickleweed)
```
Written response: 
I checked whether soil salinity and pickleweed biomass were normally distributed, since Pearson's r assumes normality. I used the Shapiro-Wilk test for both variables, and the test showed p = 0.46 for soil salinity and p = 0.11 for pickleweed biomass. Since both p-values are greater than 0.05, the data does not significantly deviate from normality, and the Pearson correlation is appropriate. 


### Running the Test
```{r}
#| label: pearson correlation
#| message: false

# run Pearson correlation between soil salinity and pickleweed biomass
cor.test(salinity$salinity_mS_cm, # predictor: soil salinity
         salinity$pickleweed, # response: pickleweed biomass
         method = "pearson") # parametric correlation test

```

## d. Results communication

To evaluate the strength of the relationship between pickleweed biomass and soil salinity, I used a Pearson's product-moment correlation. We found a moderate positive relationship between soil salinity and pickleweed biomass (Pearson’s r = 0.53, t(21) = 2.9, p = 0.01, ⍺ = 0.05), which suggests that pickleweed 
biomass tends to increase as soil salinity increases along the slough. 


## e. Test implications
You’re working on a team of people at this restoration site who are also concerned about pickleweed planting. In 2-3 sentences, write what you would communicate to them about the results of this test and what it means for pickleweed planting success at your site.

Be cognizant of your audience as you are writing: what would they need to know to take action?


## f. Double check your own work.



# Problem 2. Personal data

## a. Updating your visualizations

### Continuous variable visualization: Time duration vs Tasks completed
```{r}
#| label: continuous variable visualization
#| message: false
my_clean_data <- my_data |> 
  clean_names() |> # clean column names
  rename(duration = duration_hr_mm) |> # rename for easier use
  filter(duration != "48:00:00") |> # remove the typo entry
   # convert all durations as decimal hours (e.g. 1.5 = 1hr 30min)
   mutate(duration_decimal = case_when(
    str_detect(duration, "\\.") ~ as.numeric(duration), # if format is 4.18, keep as numeric
    str_detect(duration, ":") ~ period_to_seconds(hm(duration)) / 3600 # if format is 1:43, convert to decimal
  )) 
# base layer: ggplot
ggplot(
  data = my_clean_data,
  aes(
    x = duration_decimal, # cleaned numeric duration
    y = tasks_completed)) + # tasks completed
  # first layer: scatterplot
  geom_point(color = "forestgreen", # green points
             size = 2, 
             alpha = 0.7) + # slight transparency 
  # create readable labels, ticks every hour for clarity 
  scale_x_continuous(breaks = seq(0, 6, by = 1), # tick every hour
                     labels = c("0h", "1h", "2h", "3h", "4h", "5h", "6h")) + 
  # title plot, x-axis, y-axis
  labs(
    title = "Analyzing productivity:\nTasks Completed vs. Study Duration",
    x = "Study Duration (hours)",
    y = "Tasks Completed",
    subtitle = "Most recent observation: March 1, 2026") + # most recent date
  theme_minimal() # clean theme
```

### Categorical variable visualization:  vs Tasks completed
```{r}
#| label: categorical-variable-visualization
#| message: false
my_clean_data <- my_data |> 
  clean_names() # Make all column names lower case with underscores
ggplot(
  data = my_clean_data, # Use cleaned personal dataset
    aes(
    x = study_location, # Categorical predictor on x-axis
    y = tasks_completed,
    fill = study_location)) + # Response variable on y-axis
  geom_boxplot( # Draw a boxplot to summarize the distribution of tasks completed
    alpha = 0.6, # Make boxplots semi-transparent
    width = 0.4) + 
  geom_jitter( # Show jittered points to show individual observations
    width = 0.1, # Horizontal spread to avoid overlap when there's more data, but there is not enough data to be an issue now
    alpha = 0.7, # Semi-transparent points
    size  = 2) + # Point size
    scale_fill_manual(values = c(
    "Café" = "#FEBC11", # warm gold for café - exact match with accent
    "Library" = "#4169E1")) + # muted blue-green for library
    scale_y_continuous(breaks = c(3, 4, 5, 6, 7, 8, 9, 10, 11, 12)) + # Set specific y-axis tick marks for clarity
      labs( # Add descriptive title and axis labels
      title = "Analyzing productivity:\nTasks Completed vs. Study Location",
      x = "Study Location",
      y = "Tasks Completed",
  subtitle = "Most recent observation: March 1, 2026") +
  theme_minimal() # Set theme different from default
```

## b. Captions


# Problem 3. Affective visualization


## a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences).


## b. Create a sketch (on paper) of your idea.


## c. Make a draft of your visualization.


## d. Write an artist statement.


## e. Prep your materials to share in class.


# Problem 4. Statistical critique

## a. Revisit and summarize


## b. Visual clarity


## c. Aesthetic clarity


## d. Recommendations 









