---
title: "Homework 3"
author: "Bella Kormos"
format: html
---

[GitHub Repository](https://github.com/bellakormos/ENVS-193DS_homework-03) 

# Homework Set Up

Reading in packages:

```{r}
#| label: packages
#| message: false
library(tidyverse) # load in all packages
library(janitor)
library(here) # identifies computer's root folder \
```

Reading in data:

```{r}
#| label: reading in data
#| message: false
salinity <- read_csv(here("data", "salinity-pickleweed.csv")) # store the data and save object as "salinity" 

```

# Problem 1. Slough soil salinity

## a. An appropriate test
To determine the strength of the relationship between salinity and California pickleweed biomass, the two appropriate tests are the Pearson correlation and Spearman correlation tests. The pearson correlation (correlation coefficient, r) measures the strength between soil salinity and biomass, and assumes that variables are continuous, normally distributed, and have independent observations. The Spearman rank correlation is a non-parametric test that measures the stength of the monotonic relationship between ranked values of soil salinity and biomass, and it does not assume normality. 

## b. Create a visualization
```{r}
# base layer: ggplot
ggplot(data = salinity, # using salinity data frame
       aes(x = salinity_mS_cm, # soil salinity on x-axis (predictor)
           y = pickleweed)) + # pickleweed biomass on y-axis (response)
  # first layer: make scatter-plot 
  geom_point(color = "forestgreen",
             alpha = 0.7, # slight transparency 
             size = 2) + # slightly larger points
  # title plot, x-axis, y-axis
  labs(x = "Soil Salinity (mS/cm)", 
       y = "Pickleweed Biomass (g)",
       title = "Relationship between Soil Salinity and Pickleweed Biomass") +
  theme_light() # different from default theme
```


## c. Check your assumptions and run your test


## d. Results communication


## e. Test implications


## f. Double check your own work.



# Problem 2. Personal data

## a. Updating your visualizations


## b. Captions


# Problem 3. Affective visualization


## a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences).


## b. Create a sketch (on paper) of your idea.


## c. Make a draft of your visualization.


## d. Write an artist statement.


## e. Prep your materials to share in class.


# Problem 4. Statistical critique

## a. Revisit and summarize


## b. Visual clarity


## c. Aesthetic clarity


## d. Recommendations 









