---
title: "Homework 3"
author: "Bella Kormos"
format: html
---

[GitHub Repository](https://github.com/bellakormos/ENVS-193DS_homework-03) 

# Homework Set Up

Reading in packages:

```{r}
#| label: packages
#| message: false
library(tidyverse) # load in all packages
library(janitor)
library(here) # identifies computer's root folder \
```

Reading in data:

```{r}
#| label: reading in data
#| message: false
salinity <- read_csv("salinity-pickleweed.csv") # store the data and save object as "salinity" 
```

# Problem 1. Slough soil salinity

## a. An appropriate test


## b. Create a visualization


## c. Check your assumptions and run your test


## d. Results communication


## e. Test implications


## f. Double check your own work.



# Problem 2. Personal data

## a. Updating your visualizations


## b. Captions


# Problem 3. Affective visualization


## a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences).


## b. Create a sketch (on paper) of your idea.


## c. Make a draft of your visualization.


## d. Write an artist statement.


## e. Prep your materials to share in class.


# Problem 4. Statistical critique

## a. Revisit and summarize


## b. Visual clarity


## c. Aesthetic clarity


## d. Recommendations 









