---
title: "Homework 3"
author: "Bella Kormos"
format: html
---

[GitHub Repository](https://github.com/bellakormos/ENVS-193DS_homework-03) 
# Homework Set Up

Reading in packages:

```{r}
#| label: packages
#| message: false
library(tidyverse) # load in all packages
library(janitor)
library(here) # identifies computer's root folder
library(lubridate)

```

Reading in data:

```{r}
#| label: reading in data
#| message: false
salinity <- read_csv(here("data", "salinity-pickleweed.csv")) # store the data and save object as "salinity" 
my_data <- read_csv(here("data", "updated_personal-data.csv")) # Read in the personal data as an object named "my_data"

```

# Problem 1. Slough soil salinity

## a. An appropriate test
To determine the strength of the relationship between salinity and California pickleweed biomass, the two appropriate tests are the Pearson correlation and Spearman correlation tests. The pearson correlation (correlation coefficient, r) measures the strength between soil salinity and biomass, and assumes that variables are continuous, normally distributed, and have independent observations. The Spearman rank correlation is a non-parametric test that measures the stength of the monotonic relationship between ranked values of soil salinity and biomass, and it does not assume normality. 

## b. Create a visualization
```{r}
#| label: create visualizaiton
#| message: false
# base layer: ggplot
ggplot(data = salinity, # using salinity data frame
       aes(x = salinity_mS_cm, # soil salinity on x-axis (predictor)
           y = pickleweed)) + # pickleweed biomass on y-axis (response)
  # first layer: make scatter-plot 
  geom_point(color = "forestgreen",
             alpha = 0.7, # slight transparency 
             size = 2) + # slightly larger points
  # title plot, x-axis, y-axis
  labs(x = "Soil Salinity (mS/cm)", 
       y = "Pickleweed Biomass (g)",
       title = "Relationship between Soil Salinity and Pickleweed Biomass") +
  theme_light() # different from default theme
```


## c. Check your assumptions and run your test

### Assumption Checks

```{r}
#| label: assumption checks
#| message: false

# check normality of salinity variable using Shapiro-Wilk test
shapiro.test(salinity$salinity_mS_cm)

# check normality of biomass variable using Shapiro-Wilk test
shapiro.test(salinity$pickleweed)
```
Written response: 
I checked whether soil salinity and pickleweed biomass were normally distributed, since Pearson's r assumes normality. I used the Shapiro-Wilk test for both variables, and the test showed p = 0.46 for soil salinity and p = 0.11 for pickleweed biomass. Since both p-values are greater than 0.05, the data does not significantly deviate from normality, and the Pearson correlation is appropriate. 


### Running the Test
```{r}
#| label: pearson correlation
#| message: false

# run Pearson correlation between soil salinity and pickleweed biomass
cor.test(salinity$salinity_mS_cm, # predictor: soil salinity
         salinity$pickleweed, # response: pickleweed biomass
         method = "pearson") # parametric correlation test

```

## d. Results communication

To evaluate the strength of the relationship between pickleweed biomass and soil salinity, I used a Pearson's product-moment correlation. We found a moderate positive relationship between soil salinity and pickleweed biomass (Pearson’s r = 0.53, t(21) = 2.9, p = 0.01, ⍺ = 0.05), which suggests that pickleweed biomass tends to increase as soil salinity increases along the slough. 


## e. Test implications
Our analysis found a moderate positive relationship between soil salinity and pickleweed biomass at this restoration site, which means that in higher salinity soils, pickleweed is growing with a greater biomass. This ultimately suggests that saltier conditions are beneficial for the growth of pickleweed, which makes sense since it is a salt-tolerate species. To ensure pickleweed planting success, planting should be prioritized in higher salinity areas of the slough.


## f. Double check your own work.
```{r}
#| label: spearman tests
#| message: false

# run Spearman rank correlation as a comparison to Pearson
cor.test(salinity$salinity_mS_cm, # predictor is salinity
         salinity$pickleweed, # response is pickleweed biomass
         method = "spearman") # non-parametric rank-based correlation

```
To evaluate the strength of the relationship between pickleweed biomass and soil salinity, I used a Spearman's rank correlation rho. We found a moderate relationship between soil salinity and pickleweed biomass (Spearman p = 59, S = 824, p = 0.003, α = 0.05), which suggests that pickleweed biomass tends to increase as soil salinity increases along the slough. Both tests agree that the relationship is moderate and positive, so the result is robust with both correlation methods.  

# Problem 2. Personal data

## a. Updating your visualizations

### Continuous variable visualization: Time duration vs Tasks completed
```{r}
#| label: continuous variable visualization
#| message: false
my_clean_data <- my_data |> # use cleaned data-set
  clean_names() |> # clean column names
  rename(duration = duration_hr_mm) |> # rename for easier use
  filter(duration != "48:00:00") |> # remove the typo entry
   # convert all durations to decimal point instead of ":"
   mutate(duration_decimal = case_when(
    str_detect(duration, "\\.") ~ as.numeric(duration), # if format is 4.18, keep as numeric
    str_detect(duration, ":") ~ period_to_seconds(hm(duration)) / 3600 # if format is 1:43, convert to decimal
  )) 
# base layer: ggplot
ggplot(
  data = my_clean_data,
  aes(x = duration_decimal, # predictor var: duration
      y = tasks_completed)) + # response var: tasks completed
  # first layer: scatter-plot
  geom_point(color = "forestgreen", # green points
             size = 2, 
             alpha = 0.7) + # slight transparency 
  # create readable labels, ticks every hour for clarity 
  scale_x_continuous(breaks = seq(0, 6, by = 1), # tick every hour
                     labels = c("0h", "1h", "2h", "3h", "4h", "5h", "6h")) + 
  # title plot, x-axis, y-axis
  labs(title = "Analyzing productivity:\nTasks Completed vs. Study Duration",
       x = "Study Duration (hours)",
       y = "Tasks Completed",
    subtitle = "Most recent observation: March 1, 2026") + # most recent date
  # add clean themes and remove minor grid lines
  theme_minimal() + 
  theme(panel.grid.minor = element_blank()) 
```

### Categorical variable visualization:  vs Tasks completed
```{r}
#| label: categorical-variable-visualization
#| message: false
my_clean_data <- my_data |> 
  clean_names() # columns names lower case with underscores
# base layer: ggplot using cleaned data-set
ggplot(data = my_clean_data, 
      aes(x = study_location, # categorical predictor on x-axis
          y = tasks_completed, # response variable on y-axis
          fill = study_location)) + 
  # add a cap at whisker end for visual appeal
  stat_boxplot(
    geom = "errorbar", # adds horizontal caps at whisker ends
    width = 0.02) + # width os the caps
  # first layer: boxplot 
  geom_boxplot(alpha = 0.6, # make boxplot transparent 
               width = 0.4) + 
  # second layer: add jittered points 
  geom_jitter(width = 0.1, # horizontal spread to avoid overlap
              alpha = 0.7, # semi-transparent points
              size  = 2) + # point size
    scale_fill_manual(values = c("Café" = "#FEBC11", # warm gold for café
                                 "Library" = "#4169E1")) + # muted blue-green for library
    scale_y_continuous(breaks = seq(0, 12, by = 2), limits = c(0, 13)) + # set  y-axis tick marks every 2 tasks for clarity
    # add title and axis labels    
    labs(title = "Analyzing productivity:\nTasks Completed vs. Study Location",
           x = "Study Location",
           y = "Tasks Completed",
  subtitle = "Most recent observation: March 1, 2026") + # add subtitle with date
  # set theme, remove legend, remove minor grid lines
  theme_minimal() + 
  theme(legend.position = "none") + 
  theme(panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_blank())
```

## b. Captions
Figure 1. The relationship between study duration (hours) and tasks completed across all study sessions is represented in a scatter-plot. The most recent observation was March 1, 2026, and each green point reperesents a study session. 

Figure 2. The distribution of tasks completed by study location (Café vs Library) across study sessions is represented by the boxplot. Each black point is a study session. The library (in blue) has higher median (around 5 tasks), while Café study sessions (gold) has a smaller IQR and median. 



# Problem 3. Affective visualization

## a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences).


## b. Create a sketch (on paper) of your idea.


## c. Make a draft of your visualization.


## d. Write an artist statement.


## e. Prep your materials to share in class.


# Problem 4. Statistical critique

## a. Revisit and summarize


## b. Visual clarity


## c. Aesthetic clarity


## d. Recommendations 









